---
title: "COMPSCIX 415.2 Homework 6"
author: "Inessa Wu"
date: "March 7, 2018"
output: 
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, warning=FALSE, message=FALSE}
library (tidyverse)
library (rmarkdown)
library (mosaicData)
library (lubridate)
```

###Exercise 1

```{r}
data(Whickham)
```

####Question 1

Variables included in the data set are Outcome (Alive, Dead), Smoker (No, Yes), Age (age in years). 

####Question 2

There are 1,314 observations in this data set. Each observation represents a woman who took a survey.

####Question 3

The summary table shows that higher share of non-smokers are dead 20 years after the initial survey has been taken. This outcome does not make sense. Age has to be a factor. It's possible that larger share of non-smokers were older than that of smokers.

```{r}
Whickham %>% count(smoker, outcome)

ggplot(data=Whickham)+
  geom_bar(mapping=aes(x=smoker, fill=outcome), position="fill")
```

####Question 4

Creating a new variable age_ctg to categorize survey responders into age groups. Checking that this step has been completed correctly.

```{r}
whicham_wkng <- Whickham %>% mutate (age_ctg = case_when(age <= 44 ~ 'age <= 44', age > 44 & age <=64 ~ 'age > 44 & age <= 64', age > 64 ~ 'age > 64'))

glimpse(whicham_wkng)
```
Converting age_ctg into orgered factor and assigning values to age_fct variable and creating a visualization. Checking that this step has been completed correctly.

```{r}
whicham_wkng <- whicham_wkng %>% mutate (age_fct = factor(age_ctg, ordered=TRUE))

glimpse(whicham_wkng)

ggplot(data = whicham_wkng)+
  geom_bar(mapping=aes(x=smoker, fill=outcome), position='fill')+
  facet_grid(~age_fct)
```

Visualiziation above makes more sense. By exposing age category, now the visualization shows that a larger share of non-smokers remained alive 20 years after the survey for each age category. It also shows that share of women dead increased with age for both smokers and non-smokers.


###Exercise 2

####Question 1

Generating a random sample of size n = 10000 from a gamma(1,2) distribution and plotting a histogram or density curve

```{r}
n <- 10000

gamma_samp <- tibble(x = rgamma(n, shape = 1, scale = 2))

ggplot(gamma_samp)+
  geom_histogram(mapping=aes(x=x))

```

####Question 2
Mean of the generated sample is 1.981001 and standard deviaion is 1.973010 

```{r}
gamma_samp %>% summarize(mean_x = mean(x, na.rm = TRUE),
                      sd_x = sd(x, na.rm = TRUE)) %>%
  gather(key = stats, value = values)
```

####Question 3

Taking a sample of size n = 30 from a Gamma(1,2) distribution and then plotting the histogram. Mean and standard deviation of the generated sample are shown below:

```{r}
gamma_n30 <- gamma_samp %>% sample_n(30)

gamma_n30 %>%
  summarize(mean_x = mean(x, na.rm = TRUE),
                      sd_x = sd(x, na.rm = TRUE)) %>%
  gather(key = stats, value = values)

ggplot(gamma_n30)+
  geom_histogram(mapping=aes(x=x))
```

####Question 4

Creating a vector mean_samp that contains values of sample means for 1000 repetitions of taking a sample of size n = 30 from the Gamma(1,2) distribution and calculating sample mean.

```{r}
# create a vector with 10000 NAs

mean_samp <- rep(NA, 10000)

# start a loop
for(i in 1:10000) {
  g_samp <- rgamma(30, shape = 1, scale = 2)
  mean_samp[i] <- mean(g_samp)
}
# Convert vector to a tibble
mean_samp <- tibble(mean_samp)

```

#####Question 5

Below is a histogram of the generated collection of means from (mean_samp)

```{r}
ggplot(mean_samp)+
  geom_histogram(mapping=aes(x=mean_samp))
```

####Question 6

Calculate the mean and standard deviation of all sample means.


```{r}
mean_samp %>%
  summarize(mean_smpl = mean(mean_samp, na.rm = TRUE),
                      sd_smpl = sd(mean_samp, na.rm = TRUE)) %>%
  gather(key = stats, value = values)
```

####Question 7

None of he values are surprising. Both values are in line with the expectations based on Central Limit Theorem. Mean is close to 2, as expectged and standard deviation is close to 0.365. 

####Question  8
Repeating sampling with n=300

```{r}
# create a vector with 300 NAs

mean_samp300 <- rep(NA, 300)

# start a loop
for(i in 1:300) {
  g_samp300 <- rgamma(30, shape = 1, scale = 2)
  mean_samp300[i] <- mean(g_samp300)
}
# Convert vector to a tibble
mean_samp300 <- tibble(mean_samp300)

```

Calculating the mean and standard deviation of all sample means from a sample of n=300

```{r}
mean_samp300 %>%
  summarize(mean_smpl300 = mean(mean_samp300, na.rm = TRUE),
                      sd_smpl300 = sd(mean_samp300, na.rm = TRUE)) %>%
  gather(key = stats, value = values)
```

Repeating sampling process and with a sample size of n = 300 produced results that still match up fairly well with the theorem: mean of 2.0465739 which is close to 2 and standard deviation of 0.3892461 which is fairly close to 0.365. However, both mean and standard deviation from the sample of n=300 have values that a little further away from the Central Limit Theorem expected values that those resulting from a sample of n=10000 



