---
title: "COMPSCIX 415.2 Homework 3"
author: Inessa Wu
date: February 13, 2018
output: html_notebook
---
```{r load_packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(nycflights13)
```

###Exercise 3.8.1
1. the graph likely  does not plot all the points, can use position=jitter to see all data points
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point(position="jitter")
```
2. Amount and Factor control the  degree of jittering in geom_jitter()
```{r}
?jitter
```

3. geom_jitter() can be used with scatterplots to alleviate overplotting by giving more separations to the points that are plotted in the same space. geom_count() is used to count number of points at each location and maps the count in various size of bubbles. geom_count might be more useful in large data sets since thescatterplot may already be too cluttered. 

4. The default position adjustment for geom_boxplot() is  position = "dodge". 

```{r}
ggplot(data=mpg)+
  geom_boxplot(mapping=aes(x=drv, y=hwy))
```

###Exercise 3.9.1

2. What does labs() modifies axis titles as well as plot titles and subtitles

4. The plot of highway vs city miles per gallon shows there is a direct relationships between city and  hwy mileage performance - higher hwy  mileage corresponds with higher city mileage. 
coord_fixed()  forces a specified ratio between the physical representation of data units on the axes. It is important for clarity of data  reporesentation so that the output is not misinterpreted. In this case x and y have the same units so fixed scale will give the best potrayal of the relationship between the two. 
geom_abline() Adds reference lines; helpful for plot evaluation

###Exercise 4.4

1.The code does not work because "my_varıable" is misspelled: ı instead of i
2. Correct dota to data

```{r}
library(tidyverse)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```
Correct "fliter" to "filter"
```{r}
filter(mpg, cyl == 8)
```
 Correct "diamond" to "diamonds"
```{r}
filter(diamonds, carat > 3)
```
###Exercise 5.2.4

```{r}
library(tidyverse)
library(nycflights13)
```

1. 
Had an arrival delay of two or more hours

```{r}
filter(flights, arr_delay>=2)
```
Flew to Houston (IAH or HOU)
```{r}
filter(flights, dest=="IAH" | dest=="HOU")
```
Were operated by United, American, or Delta
```{r}
filter(flights, carrier %in% c("UA", "AA", "DL"))
```
Departed in summer (July, August, and September)

```{r}
filter(flights, month %in% c(7, 8, 9))
```

Arrived more than two hours late, but didn’t leave late
```{r}
filter(flights, dep_delay<=0, arr_delay>120)
```
Were delayed by at least an hour, but made up over 30 minutes in flight
```{r}
filter(flights, dep_delay>=60, (dep_delay-arr_delay)>30)
```
Departed between midnight and 6am (inclusive)
```{r}
filter(flights, sched_dep_time>=0000, sched_dep_time<=0600)
```

2. between() filters values within the specified range

Departed between midnight and 6am (inclusive)
```{r}
filter(flights, between(sched_dep_time, 0000, 0600))
```
3. 8,255 flights have a missing dep_time. Dep_delay, arr_time, arr_delay are slaso missing. These might be cancelled flights. 

```{r}
(missing_dep_time <- filter(flights, is.na(dep_time)))

```

```{r}
count(missing_dep_time)
```
4. 
Why is NA ^ 0 not missing? Any value to power if zero is equal to 1, including any unknown value
Why is NA | TRUE not missing? This is an OR statement which returns TRUE if either of the terms is TRUE, which is correct in this situation
Why is FALSE & NA not missing? & operator will return TRUE only if both terms are the same. In this case we do not know what NA terms are which is why FALSE is returned
Can you figure out the general rule? (NA * 0 is a tricky counterexample!). If the operation does not care whether NA is an integer or not, the response will be not missing, It will be missing if the operation can be perfomed on specific variable type only


```{r}
NA^0
```
```{r}
NA|TRUE

```
```{r}
FALSE & NA

```
###Exercise 5.4.1
1. Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights

```{r}
select(flights, dep_time, dep_delay, arr_time, arr_delay)
```
```{r}
select(flights, dep_time:arr_delay, -(sched_dep_time), -(sched_arr_time))
```

2. What happens if you include the name of a variable multiple times in a select() call. Nothing, variable listed multiple times shows up only once
```{r}
select(flights, dep_time, dep_time)
```

3. one_of() function aloows you to call a predetermined set of columns instead of explicitly listing out column headers in select() function? Might it be helpful in conjunction with this vector if the same set  of columns will be needed multiple times throughout the code.

```{r}
(vars <- c("year", "month", "day", "dep_delay", "arr_delay"))
select(flights, one_of(vars))
```
4. Argument in contians () function is case insesitive. this is surpirising given that R is case sensitive in general. Default is case insensitive. To change to case sensitive include ignore.case=FALSE

```{r}
select(flights, contains("TIME", ignore.case = FALSE))
```

