---
title: "COMPSCIX 415.2 Homework 1"
author: "Inessa Wu"
date: "January 31, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
####My Github repository for my assignments can be found at this URL: [https://github.com/InessaWu/compscix-415-2-assignments]

library(mdsr)
library(tidyverse)

```{r load_packages, warning=FALSE, message=FALSE}
```

data ("WorldCities")

####23,018 observations
####10 variables
####variables such as name, latitude, country

glimpse(WorldCities)

WorldCities <- head(WorldCities, 200)


####Extract Country column and show unique values:
country_col <- WorldCities$country

unique(country_col)

######"AD" "AE" "AF" "AG" "AI" "AL" "AM" "AO" "AR"

####Extract region column and show unique values:

region_col <- WorldCities$region

unique(region_col)

######"Europe/Andorra", "Asia/Dubai", "Asia/Kabul", "America/Antigua", "America/Anguilla",   "Europe/Tirane", Asia/Yerevan", "Africa/Luanda", "America/Argentina/Buenos_Aires", "America/Argentina/Cordoba", "America/Argentina/Salta", "America/Argentina/Tucuman", "America/Argentina/San_Juan"    

End of assignment 1 and 2