---
title: "COMPSCIX 415.2 Homework 5/Midterm"
author: "Inessa Wu"
date: "February 28, 2018"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, warning=FALSE, message=FALSE}
library(tidyverse)
library (rmarkdown)
```

###The tidyverse Packages

#### Question 1

The following packages are associated with each task below:

Plotting - ggplot2
Data munging/wrangling - dplyr
Reshaping (speading and gathering) data - tidyr
Importing/exporting data - readr

####Question 2

Below are few functions that I’ve used from each of the package listed in Question 1 for the tasks below:

Plotting - ggplot(), geom_point(), aes(), coord_flip() 
Data munging/wrangling - select(), filter(), mutate(), summarise(), arrange()
Reshaping data - gather(), spread(), separate(), unite()
Importing/exporting data (note that readRDS and saveRDS are base R functions) - read_csv(), read_delim()

###R Basics

####Question 1

This code:
My_data.name___is.too00ooLong! <- c( 1 , 2   , 3 )
should be fixed by removing the exclamation sign, as follows:

```{r}
My_data.name___is.too00ooLong <- c( 1 , 2   , 3 )
My_data.name___is.too00ooLong
```

####Question 2

This code:
my_string <- C('has', 'an', 'error', 'in', 'it)
should be fixed by changing "C"" to lower case "c" and adding ' after it, as follows:

```{r}
my_string <- c('has', 'an', 'error', 'in', 'it')
my_string
```

####Question 3

Code below outputs character/string data type, not integers data type

```{r}
my_vector <- c(1, 2, '3', '4', 5)
my_vector
```

###Data Import/Export

###Question 1

Downloading the rail_trail.txt file from Canvas (in the Midterm Exam section here) and importing it into R. 

```{r}
file_path <- 'C:/Users/ipodvorna/Documents/compscix-415-2-assignments/compscix-415-2-assignments/rail_trail.txt'
rail_trail_data <- read_delim(file = file_path, delim = "|")
```
Taking a glimpse at the imported rail_trail_data file

```{r}
glimpse(rail_trail_data)
```

####Question 2

Exporting the file into an R-specific format: 

```{r}
saveRDS(rail_trail_data, file = 'C:/Users/ipodvorna/Documents/compscix-415-2-assignments/compscix-415-2-assignments/rail_trail.rds')
```

Reload rail_trail.rds file.

```{r}
rail_trail_data_R <- readRDS('C:/Users/ipodvorna/Documents/compscix-415-2-assignments/compscix-415-2-assignments/rail_trail.rds')
```

Taking a glimpse

```{r}
glimpse(rail_trail_data_R)
```

