---
title: "COMPSCIX 415.2 Homework 5/Midterm"
author: "Inessa Wu"
date: "February 28, 2018"
output: 
  html_document:
    toc: true
    toc_float: true
---


####GitHub repository: https://github.com/InessaWu/compscix-415-2-assignments


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, warning=FALSE, message=FALSE}
library(tidyverse)
library (rmarkdown)
```

###The tidyverse Packages

####Question 1

The following packages are associated with each task below:

Plotting - ggplot2
Data munging/wrangling - dplyr
Reshaping (speading and gathering) data - tidyr
Importing/exporting data - readr

####Question 2

Below are few functions that I’ve used from each of the package listed in Question 1 for the tasks below:
 
Plotting - ggplot(), geom_point(), aes(), coord_flip() 
Data munging/wrangling - select(), filter(), mutate(), summarise(), arrange()
Reshaping data - gather(), spread(), separate(), unite()
Importing/exporting data (note that readRDS and saveRDS are base R functions) - read_csv(), read_delim()

###R Basics

####Question 1

This code:
My_data.name___is.too00ooLong! <- c( 1 , 2   , 3 )
should be fixed by removing the exclamation sign, as follows:

```{r}
My_data.name___is.too00ooLong <- c( 1 , 2   , 3 )
My_data.name___is.too00ooLong
```

####Question 2

This code:
my_string <- C('has', 'an', 'error', 'in', 'it)
should be fixed by changing "C"" to lower case "c" and adding ' after it, as follows:

```{r}
my_string <- c('has', 'an', 'error', 'in', 'it')
my_string
```

####Question 3

Code below outputs character/string data type, not integers data type

```{r}
my_vector <- c(1, 2, '3', '4', 5)
my_vector
```

###Data Import/Export

####Question 1

Downloading the rail_trail.txt file from Canvas (in the Midterm Exam section here) and importing it into R. 

```{r}
file_path <- 'C:/Users/ipodvorna/Documents/compscix-415-2-assignments/compscix-415-2-assignments/rail_trail.txt'
rail_trail_data <- read_delim(file = file_path, delim = "|")
```
Taking a glimpse at the imported rail_trail_data file

```{r}
glimpse(rail_trail_data)
```

####Question 2

Exporting the file into an R-specific format: 

```{r}
saveRDS(rail_trail_data, file = 'C:/Users/ipodvorna/Documents/compscix-415-2-assignments/compscix-415-2-assignments/rail_trail.rds')
```

Reload rail_trail.rds file.

```{r}
rail_trail_data_R <- readRDS('C:/Users/ipodvorna/Documents/compscix-415-2-assignments/compscix-415-2-assignments/rail_trail.rds')
```

Taking a glimpse

```{r}
glimpse(rail_trail_data_R)
```

###Visualization

####Question 1.

The following things are wrong with the graph:

1. For each category of the graph, percentage of respondents shown answering "yes" or "no" do not add up to 100%, e.g. for "under 45" category answering "yes" or "no" adds up to 95%. The same problem exists across all shown categories

2. The relative sizes of the bubbles are not of the same scale with smaller values specifically made to apper smaller than they shoudl have been

3. Chart does not show how large each of the age categories are relative to each other or as % of 100% which makes it difficult to interpret correctly. May be all women were under 45 and men where mostly over 45 years old..

####Question 2

Reproducing graphic using the diamonds data set:

```{r}
ggplot(data=diamonds)+
  geom_boxplot(mapping = aes(x=(reorder(cut,-carat, FUN=mean)), y=carat, fill=color), position="identity")+
  labs(x="CUT OF DIAMOND", y="CARAT OF DIAMOND")+
  coord_flip()
```

####Question 3

By chaging position to "dodge", graph now shows separate box plots for each color of diamond within each cut category instead oflayering them

```{r}
ggplot(data=diamonds)+
  geom_boxplot(mapping = aes(x=(reorder(cut,-carat, FUN=mean)), y=carat, fill=color), position="dodge")+
  labs(x="CUT OF DIAMOND", y="CARAT OF DIAMOND")+
  coord_flip()
```

###Data munging and wrangling

####Question 1

table2 dataset is not tidy.It contains "cases" and "population" variables combined in the same column under "type" header. The data set needs to be spread to seprate "cases" and "population" variables.

```{r}
table2_tidy <-  
  spread(table2, type, count)

print(table2_tidy)
```

####Question 2

Creating a new column in the diamonds data set called price_per_carat that shows the price of each diamond per carat. Only code is shown, not the output

```{r include=FALSE}
mutate(diamonds, price_per_carat=price/carat)
```

####Question 3

Code for getting number of diamonds, and the applicable proportion with price > 10000 and a carat < 1.5

```{r}
diamonds %>% summarise(number_of_diamonds = sum(price > 10000 & carat < 1.5),
            prop_diamonds = mean(price > 10000 & carat < 1.5))
  
```

The results make sense, there is onlu 1.5% of diamonds that are very high priced (greater than $10K) while under 1.5 carats in size. The output does reveal why this might be happening. This is mostl likely driven by the quality of the diamonds such as color, cut, and calrity


###EDA

####Question 1

txhousing data set covers data from January 2000 to July 2015

```{r}
txhousing %>%
  unite(year_month, year, month) %>%
  summarize (min(year_month), max(year_month), min(date), max(date))

```

####Question 2

txhousing dataset has 46 cities represented

```{r}
txhousing %>%
  summarise( count_dist=n_distinct(city))
        
```

####Question 3

Houston in July of 2015 had the highest number of sales

```{r}
txhousing %>% 
  unite(year_month, year, month) %>%
  group_by(city, year_month) %>%
    summarize (number_sales = sum(sales)) %>%
  arrange(desc(number_sales))

```

####Question 4

There is a linear relationships between sales and listings (values are correlated) - Sales increase with the increase in number of listings.

```{r}
txhousing %>% ggplot()+
geom_point (mapping=aes(x=listings, y=sales))
```

```{r}
txhousing %>% 
  ggplot()+
  geom_bin2d(mapping=aes(x=listings, y=sales))
```

####Question 5

Sales are missing for 568 records (6.6%). Detail by city is shown below with South Padre Island having the largest number proportion of sales records missing at 62%

```{r}
txhousing %>% 
  summarize (sales_missing = sum(is.na(sales)), 
             prop_missing = mean(is.na(sales)))
```


```{r}
txhousing %>% 
  group_by(city) %>%
  summarize (sales_missing = sum(is.na(sales)), 
             prop_missing = mean(is.na(sales))) %>%
  arrange (desc(prop_missing))
```

####Question 6

Distribution of the median sales price when grouped by city differ, as shown on the boxplot below. City with large number of outliers, such as Dallas, Houston, and COllin County would warrant more investigation. 

Filter of greater than 500 sales is useful to cut the noise and be able to clearly look at and interpret the results.

```{r}
txhousing %>% 
  unite(year_month, year, month, remove=FALSE) %>%
  group_by(year_month) %>%
  filter (sales > 500) %>%
 
  ggplot()+
  geom_boxplot(mapping=aes(x=reorder(city, median, FUN=median.default), y=median))+
  labs(x="City")+
  coord_flip()
```




